---
params:
  district: 792
  district_name: "Shelby County"
  report_date: !r format(lubridate::today(), "%B %d, %Y")
  data_file_date: !r lubridate::today()
  teacher_file_date: !r lubridate::today()
  teacher_file_month: "-2019-11"
title: "`r paste('Monthly DPSIG/SLIG Data Report for ', params$district_name)`"
date: "`r params$report_date`"
output:
  pdf_document:
    latex_engine: xelatex
classoption: landscape
fontsize: 12pt
mainfont: Arial
---

```{r setup, include=FALSE}

library(knitr)
library(ggpubr)
library(magrittr)
library(openxlsx)
library(tidyverse)

directory_onedrive <- "C:/Users/CA20397/OneDrive - TN Dept of Education/"
directory_project <- "goals/dsi-grant-monitoring/2020-dpsig-slig-monitoring/"

opts_chunk$set(
  echo = F,
  fig.height = 7,
  fig.width = 9,
  message = F,
  results = "hide",
  warning = F
)

opts_knit$set(root.dir = str_c(directory_onedrive, directory_project))

```

This monthly report provides an update on chronic absenteeism and student discipline in your district's CSI school(s) as well as CSI schools in other districts. With some exceptions, CSI schools (short for Comprehensive Support and Improvement) are also known as Priority schools.

The Division of School Improvement (DSI) will refer to this report in regular check-ins with districts that have CSI schools, and DSI will also ask the following questions to learn supplement this report with additional information:

* Student chronic absenteeism
    + What factors have contributed to students' chronic absenteeism this month (or the lack thereof)?  How do you plan (or have you planned) to address these factors?
    + To what extent has chronic absenteeism varied by student group (e.g., Black or African American, English Learners, Male, Students with Disabilities)?

* Student out-of-school discipline
    + What factors have contributed to students' discipline rates this month (or the lack thereof)? How do you plan (or have you planned) to address these factors?
    + To what extent have discipline rates varied by student group (e.g., Black or African American, English Learners, Male, Students with Disabilities)?

*	Student mobility
    + What factors have contributed to student mobility this month (or the lack thereof)? How do you plan (or have you planned) to address these factors?
    + To what extent has student mobility varied by student group (e.g., Black or African American, English Learners, Male, Students with Disabilities)?
    + Where do students who leave your Priority schools typically go? Where do new students typically come from?

\newpage

*	Teachers
    + What factors have contributed to teacher turnover this month (or the lack thereof)? How do you plan (or have you planned) to address these factors?
    + What factors have contributed to teacher chronic absenteeism this month (or the lack thereof)? How do you plan (or have you planned) to address these factors?

Disclaimer: No graph in this report can predict end-of-year outcomes with 100% accuracy due to a variety of reasons (e.g., changes in business rules, data quality throughout the year). Draw conclusions cautiously.
\newpage

```{r data, include=FALSE}

absence_district_month <-
  read_csv(
    str_c("data/absence-district-month-",
          params$data_file_date,
          ".csv")
  )

absence_school <-
  read_csv(
    str_c("data/absence-school-",
          params$data_file_date,
          ".csv")
  )

discipline_district_month <-
  read_csv(
    str_c("data/discipline-district-month-",
          params$data_file_date,
          ".csv")
  )

discipline_school <-
  read_csv(
    str_c("data/discipline-school-",
          params$data_file_date,
          ".csv")
  )

mobility_district_month <-
  read_csv(
    str_c("data/mobility-district-month-",
          params$data_file_date,
          ".csv")
  )

mobility_school <-
  read_csv(
    str_c("data/mobility-school-",
          params$data_file_date,
          ".csv")
  )

month_levels <-
  c("Aug", "Sep", "Oct", "Nov", "Dec",
    "Jan", "Feb", "Mar", "Apr", "May", "Jun")

schools_csi <- read_csv("data/schools-csi.csv")

teacher_school <-
  read_csv(
    str_c("data/teacher-school-",
          params$data_file_date,
          ".csv")
  )

```

```{r graph_line_district, include=FALSE}

temp <-
  list(
    absence_district_month,
    discipline_district_month,
    mobility_district_month,
    mobility_district_month
  ) %>%
  map(
    ~ .x %>%
      filter(district == params$district) %>%
      mutate_at(
        "calendar_month",
        funs(factor(., levels = c(8:12, 1:6), labels = month_levels))
      )
  )

graph_line_district <-
  pmap(
    .l = list(
      data_frames = temp,
      y_variables_quo = quos(
        pct_students_chr_absent,
        pct_students_disciplined,
        pct_students_entered,
        pct_students_exited
      ),
      y_variables_chr = c(
        "pct_students_chr_absent",
        "pct_students_disciplined",
        "pct_students_entered",
        "pct_students_exited"
      ),
      y_labels = c(
        "Chronically Absent",
        "Disciplined",
        "Entered",
        "Exited"
      ),
      limits = list(
        c(0, 100),
        c(0, 50),
        c(0, 100),
        c(0, 100)
      ),
      colors = list(
        c("skyblue", "red"),
        c("skyblue", "orange"),
        "purple",
        "deeppink3"
      )
    ),
    .f = ~ ..1 %>%
      ggplot(aes(x = calendar_month, y = !! ..2)) +
      geom_line(
        data = ..1 %>% filter(!str_detect(measure_name, "AMO")),
        aes(color = measure_name, group = measure_name),
        size = 1
      ) +
      geom_line(
        data = ..1 %>% filter(str_detect(measure_name, "AMO")),
        aes(color = measure_name, group = measure_name),
        linetype = "dashed", size = 1
      ) +
      geom_point(aes(color = measure_name), size = 3) +
      scale_y_continuous(
        name = str_c("Percent of Students ", ..4),
        limits = c(0, pmin(100, 1.25 * max(..1[[..3]], na.rm = T)))
      ) +
      scale_color_manual(values = c("gray", ..6)) +
      theme_minimal() +
      theme(
        axis.title.x = element_blank(),
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
      ) +
      ggtitle("Your CSI Schools This Year vs. Last Year")
  ) %>%
  set_names(nm = c("absence", "discipline", "mobility_enter", "mobility_exit"))

```

```{r graph_bar_school, include=FALSE}

graph_bar_school <-
  pmap(
    .l = list(
      data_frames = list(
        absence_school,
        discipline_school,
        mobility_school,
        mobility_school
      ),
      y_variables_quo = quos(
        pct_students_chr_absent_2020_ytd,
        pct_students_disciplined_2020_ytd,
        pct_students_entered_2020_ytd,
        pct_students_exited_2020_ytd
      ),
      y_labels = c(
        "Chronically Absent",
        "Disciplined",
        "Entered",
        "Exited"
      ),
      fill_colors = c("red", "orange", "purple", "deeppink3")
    ),
    .f = ~ ..1 %>%
      left_join(schools_csi, by = c("district", "school")) %>%
      mutate(
        school_name = reorder(school_name, !! ..2, max),
        in_district = district == params$district,
        y_limit = max(!! ..2, na.rm = T)
      ) %>%
      mutate_at(vars(!! ..2), funs(if_else(. < 0.5 | is.na(.), 0.5, .))) %>%
      ggplot(aes(x = school_name, y = !! ..2)) +
      geom_col(aes(fill = in_district)) +
      scale_x_discrete(name = "Each bar represents one CSI school.") +
      scale_y_continuous(
        name = str_c("Percent of Students ", ..3),
        limits = c(0, pmin(100, 1.25 * max(..1[[quo_name(..2)]], na.rm = T)))
      ) +
      scale_fill_manual(
        name = "School in your district?",
        labels = c("No", "Yes"),
        values = c("gray", ..4)
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top"
      ) +
      ggtitle("Your CSI Schools Compared to Other Districts' This Year")
  ) %>%
  set_names(nm = c("absence", "discipline", "mobility_enter", "mobility_exit"))

```

```{r table_school, include=FALSE}

# List names

list_names <- c("absence", "discipline", "mobility", "teacher_exit_ela_math")

# Captions

fill_descriptions <- list(
  c("green", "red"),
  c("green", "red"),
  c("less darkly", "more darkly"),
  c("less darkly", "more darkly")
)

outcome_names <- c(
  "Chronic Absenteeism",
  "Out-of-School Discipline Rate",
  "Student Mobility",
  "ELA and Math Teacher Exits"
)

captions <-
  map2(
    map(
      fill_descriptions,
      ~ str_c("School ranks show where schools stand among all 92 CSI schools ",
              "(1 corresponds to the lowest percent and is shaded ", .x[[1]], ", 92 corresponds to the ",
              "highest and is shaded ", .x[[2]], "). ")
    ),
    outcome_names %>%
      map(
        ~ str_c("The 2018-19 Year-to-Date column ",
                "shows each school's ", .x, " at this time last year.")
      ) %>%
      map_if(~ str_detect(.x, "Teacher Exits"), ~ ""),
    .f = function(a, b) str_wrap(str_c(a, b), width = 100)
  )

# Other parameters for map() calls below

columns <-
  list(
    levels_absence = c(
      "pct_students_chr_absent_2019_eoy",
      "pct_students_chr_absent_2019_ytd",
      "pct_students_chr_absent_2020_ytd",
      "school_rank_2019_eoy",
      "school_rank_2020_ytd"
    ),
    levels_discipline = c(
      "pct_students_disciplined_2019_eoy",
      "pct_students_disciplined_2019_ytd",
      "pct_students_disciplined_2020_ytd",
      "school_rank_2019_eoy",
      "school_rank_2020_ytd"
    ),
    levels_mobility = c(
      "pct_students_entered_2019_ytd",
      "pct_students_entered_2020_ytd",
      "school_rank_entered_2020_ytd",
      "pct_students_exited_2019_ytd",
      "pct_students_exited_2020_ytd",
      "school_rank_exited_2020_ytd"
    ),
    levels_teacher_exit_ela_math = c(
      "n_exited_ela",
      "pct_exited_ela",
      "school_rank_exited_ela",
      "n_exited_math",
      "pct_exited_math",
      "school_rank_exited_math"
    ),
    labels_absence_discipline = c(
      "2018-19 End-of-Year",
      "2018-19 Year-to-Date",
      "2019-20 Year-to-Date",
      "2018-19 End-of-Year Rank",
      "2019-20 Year-to-Date Rank"
    ),
    labels_mobility = c(
      "2018-19 Year-to-Date Entrance Rate",
      "2019-20 Year-to-Date Entrance Rate",
      "2019-20 Year-to-Date Rank (Entrance Rate)",
      "2018-19 Year-to-Date Exit Rate",
      "2019-20 Year-to-Date Exit Rate",
      "2019-20 Year-to-Date Rank (Exit Rate)"
    ),
    labels_teacher_exit_ela_math = c(
      "# ELA Teachers Exited Year-to-Date",
      "% ELA Teachers Exited Year-to-Date",
      "2019-20 Year-to-Date Rank (ELA %)",
      "# Math Teachers Exited Year-to-Date",
      "% Math Teachers Exited Year-to-Date",
      "2019-20 Year-to-Date Rank (Math %)"
    )
  )

variables_reverse_fill <- c(
  "school_rank_exited_2020_ytd", # student mobility
  "school_rank_exited_math" # teacher mobility
)

# Data frames

temp <-
  pmap(
    .l = list(
      data_frames = list(
        absence_school,
        discipline_school,
        mobility_school,
        teacher_school
      ),
      column_levels = list(
        columns$levels_absence,
        columns$levels_discipline,
        columns$levels_mobility,
        columns$levels_teacher_exit_ela_math
      ),
      column_labels = list(
        columns$labels_absence_discipline,
        columns$labels_absence_discipline,
        columns$labels_mobility,
        columns$labels_teacher_exit_ela_math
      )
    ),
    .f = ~ ..1 %>%
      filter(district == params$district) %>%
      left_join(schools_csi, by = c("district", "school")) %>%
      select(district, district_name, school, school_name, !!! ..2) %>%
      gather(
        column_name, cell_value,
        -district, -district_name, -school, -school_name
      ) %>%
      mutate(
        cell_label = if_else(
          str_detect(column_name, "pct"),
          str_c(round(cell_value), "%"),
          as.character(cell_value)
        ),
        fill_var = case_when(
          column_name %in% variables_reverse_fill ~ -1 * cell_value,
          str_detect(column_name, "rank") ~ cell_value
        )
      ) %>%
      mutate_at(
        "column_name",
        funs(factor(., levels = ..2, labels = ..3))
      ) %>%
      mutate_at(
        "school_name",
        funs(factor(., levels = base::sort(unique(.), decreasing = T)))
      )
  )

# Tables (ggplot objects)

table_school <-
  pmap(
    .l = list(
      data_frames = temp,
      outcome_names = outcome_names,
      fill_colors_low = c(rep("darkseagreen", 2), "deeppink3", "cyan3"),
      fill_colors_high = c(rep("tomato", 2), "purple", "darkorange2"),
      limits = list(
        c(1, nrow(schools_csi)),
        c(1, nrow(schools_csi)),
        c(-1 * nrow(schools_csi), nrow(schools_csi)),
        c(-1 * nrow(schools_csi), nrow(schools_csi))
      ),
      midpoints = c(rep(median(1:nrow(schools_csi)), 2), 0, 0),
      captions = captions,
      font_sizes = c(5, 5, 5, 4)
    ),
    .f = ~ ..1 %>%
      ggplot(aes(x = school_name, y = column_name)) +
      geom_tile(
        aes(fill = fill_var), show.legend = F,
        color = "black"
      ) +
      geom_text(
        data = ..1 %>% filter(!str_detect(column_name, "Rank")),
        aes(label = cell_label), hjust = 1, nudge_y = 0.2, size = ..8
      ) +
      geom_text(
        data = ..1 %>% filter(str_detect(column_name, "Rank")),
        aes(label = cell_label), size = ..8
      ) +
      coord_flip() +
      scale_y_discrete(
        labels = label_wrap_gen(width = 10),
        position = "right"
      ) +
      scale_fill_gradient2(
        low = ..3, mid = "gray85", high = ..4,
        limits = ..5,
        midpoint = ..6,
        na.value = "white"
      ) +
      theme_minimal() +
      theme(
        axis.title = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.grid = element_blank()
      ) +
      labs(
        title = str_c(..2, " in Your CSI Schools"),
        caption = ..7
      )
  ) %>%
  set_names(nm = list_names)

```

# Student Chronic Absenteeism

This outcome, Percent of Students Chronically Absent, is calculated as follows: (number of students who have missed at least 10% of school days) divided by (number of students enrolled for at least half the school year).

The graph below shows year-to-date chronic absenteeism data *pooled across all of your district's CSI schools*, measured at the end of each calendar month.

* The gray line represents last year.
* The blue line represents Annual Measureable Objectives (AMOs) for this year.
* The red line represents actual data this year.

```{r graph_line_district_absence, fig.height=4}
graph_line_district$absence
```
\newpage

The graph below shows year-to-date chronic absenteeism data *for each CSI school individually*. For more detail see the table on the following page.

* Each red bar is one of your district's CSI schools.
* Each gray bar is a CSI school outside your district.

```{r graph_bar_school_absence, fig.height=5}
graph_bar_school$absence
```
\newpage

```{r table_school_absence}
table_school$absence
```
\newpage

# Out-of-School Discipline Rates

This outcome, Percent of Students Disciplined, is calculated as follows: (number of students with at least 1 OSS or expulsion) divided by (number of students enrolled at any point this year).

The graph below shows year-to-date out-of-school discipline data *pooled across all of your district's CSI schools*, measured at the end of each calendar month.

* The gray line represents last year.
* The blue line represents Annual Measureable Objectives (AMOs) for this year. (Note: Discipline AMOs are not actually part of state accountability policy. These AMOs here are for informational purposes only.)
* The orange line represents actual data this year.

```{r graph_line_district_discipline, fig.height=4}
graph_line_district$discipline
```
\newpage

The graph below shows year-to-date out-of-school discipline data *for each CSI school individually*. For more detail see the table on the following page.

* Each orange bar is one of your district's CSI schools.
* Each gray bar is a CSI school outside your district.

```{r graph_bar_school_discipline, fig.height=5}
graph_bar_school$discipline
```
\newpage

```{r table_school_discipline}
table_school$discipline
```
\newpage

# Student Mobility

The first outcome here, Percent of Students Entered, is calculated as follows: (number of students with an enrollment date after the first day of school) divided by (number of students enrolled at any point this year).

The second outcome here, Percent of Students Exited, is calculated as follows: (number of students with a withdrawal date) divided by (number of students enrolled at any point this year).

The graphs below show year-to-date student mobility data *pooled across all of your district's CSI schools*, measured at the end of each calendar month.

* The gray lines represent last year.
* The purple and pink lines, respectively, represent actual data this year.

```{r graph_line_district_mobility, fig.height=4}
ggarrange(
  graph_line_district$mobility_enter +
    labs(subtitle = "Student Entrance Rate"),
  graph_line_district$mobility_exit +
    labs(subtitle = "Student Exit Rate"),
  ncol = 2, nrow = 1
)
```
\newpage

The graphs below show year-to-date student mobility data *for each CSI school individually*. For more detail see the table on the following page.

* Each purple or pink bar is one of your district's CSI schools.
* Each gray bar is a CSI school outside your district.
* Note that schools are not necessarily in the same order in both graphs.

```{r graph_bar_school_mobility, fig.height=5}
ggarrange(
  graph_bar_school$mobility_enter +
    labs(subtitle = "Student Entrance Rate"),
  graph_bar_school$mobility_exit +
    labs(subtitle = "Student Exit Rate"),
  ncol = 1, nrow = 2
)
```
\newpage

```{r table_school_mobility}
table_school$mobility
```
\newpage

# Teacher Mobility

This outcome, ELA and Math Teacher Exits, is calculated as follows: ([number of teachers start-of-year] minus [number of teachers persisted this year]) divided by (number of teachers start-of-year). These calculations are done separately for ELA and Math from the data submitted by your district.

```{r table_school_teacher_exit_ela_math, fig.height=5.5}
table_school$teacher_exit_ela_math
```
